On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions

### 2013-09-30 paper2.tex
paper 2 corrected

%\newcommand{\rmn}{\mathrm}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\expval}[1]{\left\langle #1 \right\rangle}
\newcommand{\RA}[3]{#1$^{\mathrm{h}}$#2$^{\mathrm{m}}$#3$^{\mathrm{s}}$}
\newcommand{\Dec}[3]{#1$^{\circ}$#2\arcmin#3\arcsec}
\newcommand{\dps}{\displaystyle}
\newcommand{\eps}{\varepsilon}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vel}{\upsilon}
%
\usepackage{graphicx}
\usepackage{natbib}
%
\usepackage{txfonts}

@@

%MNRAS
\title[On the Physics of Radio Halos in Galaxy Clusters]{On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2}$, Christoph Pfrommer$^{3}$ and Francisco Prada$^{4,5,1}$\\
$^{1}$GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands, f.zandanel@uva.nl\\
$^{2}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany, christoph.pfrommer@h-its.org\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf The underlying physics of giant and mini radio halos in galaxy clusters
    is still an open question. We find that mini halos (such as in Perseus and
    Ophiuchus) can be explained by radio-emitting electrons that are generated
    in hadronic cosmic ray (CR) interactions with protons of the intra-cluster
    medium. By contrast, the hadronic model either fails to explain the extended
    emission of giant radio halos (as in Coma at low frequencies) or would
    require a flat CR profile, which can be realized through outward streaming
    and diffusion of CRs (in Coma and A2163 at 1.4~GHz).  We suggest that a
    second, leptonic component could be responsible for the missing flux in
    giant halos within a new hybrid scenario and we describe its possible
    observational consequences.  To study the hadronic emission component of the
    radio halo population statistically, we use a cosmological mock galaxy
    cluster catalog built from the MultiDark simulation. Because of the
    properties of CR streaming and the different scalings of the X-ray
    luminosity ($L_{\rmn{X}}$) and the Sunyaev-Zel'dovich flux ($Y$) with gas
    density, our model can simultaneously reproduce the observed bimodality of
    radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
    unimodal distribution of radio-halo luminosity versus $Y$; thereby
    suggesting a physical solution to this apparent contradiction. We predict
    radio halo emission down to the mass scale of galaxy groups, which
    highlights the unique prospects for low-frequency radio surveys (such as the
    LOFAR Tier 1 survey) to increase the number of detected radio halos by at
    least an order of magnitude.}
\end{abstract}

@@

{\bf The presence of large-scale diffuse radio synchrotron emission in clusters of
galaxies proves the existence of relativistic electrons and magnetic fields
permeating the intra-cluster medium (ICM). Diffuse cluster radio emission can be
separated in two classes: radio halos and radio relics (see, e.g., 
\citealp{2012A&ARv..20...54F}).  Radio \mbox{(mini-)}halos (RHs) are centered 
on clusters and characterized by a regular and unpolarized morphology, resembling 
the morphology of the thermal X-ray emission.}

Two principal models have been proposed to explain RHs.  In the ``hadronic
model'' the radio emitting electrons are produced in hadronic cosmic ray (CR)
proton interactions with protons of the ambient thermal ICM, requiring only a
very modest fraction of (at most) a few percent of CR-to-thermal pressure
(\citealp{1980ApJ...239L..93D,1982AJ.....87.1266V, 1999APh....12..169B,
  2000A&A...362..151D, 2001ApJ...562..233M,2001ApJ...559...59M,
  2003MNRAS.342.1009M,2003A&A...407L..73P, 2004A&A...413...17P,
  2004MNRAS.352...76P, 2007IJMPA..22..681B, 2008MNRAS.385.1211P,
  2008MNRAS.385.1242P, 2009JCAP...09..024K, 2010MNRAS.401...47D,
  2010arXiv1003.0336D, 2010arXiv1003.1133K, 2010arXiv1011.0729K,
  2011A&A...527A..99E}).  CR protons and heavier nuclei, like electrons, can be
accelerated and injected into the ICM by structure formation shocks, active
galactic nuclei (AGN) and galactic winds.  Due to their higher masses with
respect to electrons, CR protons are accelerated more efficiently to
relativistic energies and are expected to show a ratio of the spectral energy
flux of CR protons to electrons above 1 GeV of about 100, similarly to what is
observed in our Galaxy \citep{2002cra..book.....S}. Additionally, CR protons
have a radiative cooling time larger than that of the electrons by the square of
the mass ratio and therefore can accumulate in clusters for cosmological times
\citep{1996SSRv...75..279V}. Indeed, CR electrons suffer more severe energy
losses via synchrotron and inverse Compton emission at particle energies $E
\gtrsim 100$~MeV, and Coulomb losses below that energy range.

@@

RHs can be divided in two classes. Giant radio halos are typically associated
with merging clusters and have large extensions, e.g., the Coma RH has
an extension of about 2~Mpc. Radio mini-halos are associated with relaxed
clusters that harbor a cool core and typically extend over a few hundred
kilo-parsecs, e.g., the Perseus radio mini-halo has an extension of about
0.2~Mpc.  The observed morphological similarities with the thermal X-ray
emission suggests that RHs may be of hadronic origin. In fact, cool-core
clusters (CCCs) are characterized by high thermal X-ray emissivities and ICM
densities that are more peaked in comparison to (merging) non cool-core clusters
(NCCCs) (e.g., \citealp{2008A&A...487..431C}). This distinct difference in the
ICM density structure of CCCs and NCCCs would be reflected in the morphology 
of the two observed classes of RHs.

The RH luminosity seems to be correlated to the thermal X-ray luminosity (e.g.,
\citealp{2009A&A...507..661B,2011A&A...527A..99E}). However, a large fraction
of clusters does not exhibit significant diffuse synchrotron emission at current
sensitivity limits. Instead, stacking subsamples of luminous X-ray clusters
reveals a signal of extended diffuse radio emission that is below
the radio upper limits on individual clusters \citep{2011ApJ...740L..28B}
suggesting that at least a subset of apparently ``radio-quiet'' clusters shows a
low-level diffuse emission. Galaxy clusters with the same thermal X-ray
luminosity show an apparent bimodality with respect to their radio
luminosity. 
%5Either they harbor a RH or they do not have any detectable diffuse
%radio emission \citep{2009A&A...507..661B,2011A&A...527A..99E}. 
This suggests the existence of a switch-on/switch-off mechanism that is able to change the
radio luminosity by more than one order of magnitude.  While such a mechanism
could be easily realized in the framework of the re-acceleration model
\citep{2009A&A...507..661B}, the \emph{classical} hadronic model predicts the
presence of RHs in all clusters. The failure to reproduce the observed cluster
radio-to-X-ray bimodality was one of the main criticisms against the hadronic
model.  Additionally, the classical hadronic model cannot reproduce some
spectral features observed in clusters, such as the total spectral (convex)
curvature claimed in the Coma RH or the spectral steepening observed at
the boundary of some RHs. 
%However, the recent report of spectral flattening with
%frequency of the RH in A2256 \citep{2012A&A...543A..43V} could easily be
%accommodated in the hadronic model, which naturally produces such a concave
%spectrum \citep{2010MNRAS.409..449P}. 
%This raises the interesting question
%whether such a variability among different sources that are generally classified
%as `radio halos' signals the presence of richer underlying physics---a question
%that we will address in this paper.

\cite{2011A&A...527A..99E} tried to asses these problems of the classical hadronic 
model by analyzing CR transport processes within a cluster. While CR advection with 
turbulently driven flows results in centrally enhanced CR profiles, the propagation in 
form of CR streaming and diffusion produces a flattening of CR profiles. Hence, 
different CR transport phenomena may also account for the observed bimodality 
of the radio luminosity in the hadronic model and may also explain the spectral 
features observed in some clusters. {\bf This has been recently confirmed 
by \cite{2013arXiv1303.4746W}. In particular, they show that when considering turbulent
damping, CRs can stream at super-Alfvenic velocities.} 
Note that these phenomena were not considered in earlier analytical works 
(e.g., \citealp{2004A&A...413...17P}) as well as in previous hydrodynamic 
simulations (e.g., \citealp{2001ApJ...562..233M, 
2008MNRAS.385.1211P, 2010MNRAS.409..449P}).
{\bf A satisfactory theory of CR transport in clusters does not exist yet, 
however CR streaming and diffusion may represent an intriguing solution
for the issues of the classical hadronic model.}

@@

An observational test that is able to disentangle between the hadronic and
re-acceleration models is the gamma-ray emission resulting from neutral pion
decays, a secondary product of the hadronic CR interaction with protons of the
ICM, which is not predicted by the re-acceleration model. Such observational
efforts have been undertaken in the last few years (for space-based cluster
observations in the GeV-band, see \citealt{2003ApJ...588..155R,
  2010JCAP...05..025A,2010ApJ...717L..71A,2012arXiv1207.6749H,
  2013arXiv1308.5654T,2013arXiv1308.6278H,2013arXiv1309.0197P}; for ground-based
observations in the energy band $\gtrsim100$~GeV, see
\citealt{2006ApJ...644..148P, 2008AIPC.1085..569P,
  2009arXiv0907.0727T,2009A&A...495...27A, 2009arXiv0907.3001D,
  2009arXiv0907.5000G, cangaroo_clusters, 2009ApJ...706L.275A,
  2010ApJ...710..634A, 2011arXiv1111.5544M,2012...VERITAS,2012A&A...545A.103H})
without being able to detect cluster gamma-ray emission. Current gamma-ray 
limits enable us to constrain the average CR-to-thermal pressure to be less than a few
percent, and the maximum CR acceleration efficiency at structure formation shocks 
to be $<50\%$. Alternatively, this may indeed suggest the presence of non-negligible 
CR transport processes into the outer cluster regions. 
%significant constraints on the relative CR pressure contribution. 
%The long observational campaign of the Perseus galaxy cluster
%performed by the Major Atmospheric Gamma-ray Cherenkov (MAGIC) telescopes
%constrains the cluster average CR-to-thermal pressure to be less than a few
%percent \citep{2010ApJ...710..634A,2011arXiv1111.5544M}. Comparing the
%corresponding upper limits to predictions of simulations by
%\cite{2010MNRAS.409..449P} constrains the maximum CR acceleration efficiency at
%structure formation shocks to be $<50\%$.  Alternatively, this may indeed
%suggest the presence of non-negligible CR transport processes into the outer
%cluster regions as suggested by \cite{2011A&A...527A..99E}.  Constraints at the
%same level are confirmed by the \emph{Fermi}-Large Area Telescope (LAT) data on
%the Coma cluster \citep{2011arXiv1105.3240P,2012AAS...21920701Z,
%2012...VERITAS,2012arXiv1207.6749H}. To summarize, gamma-ray observations are 
%not yet sensitive enough to severely challenge hadronic models of RHs.

@@

{\bf The main scope of this work is to upgrade the classical hadronic model with
the inclusion of possible CR transport phenomena, and to make predictions for
future radio surveys. The outline is as follows. In Section~\ref{sec:2.3}, we 
construct a model for the CR proton distribution in clusters that merges results 
of hydrodynamic cluster simulations and an analytical model for microscopic 
CR transport processes. In Section~\ref{sec:3}, we model observed surface 
brightness profiles of individual RHs within the hadronic scenario and explore the 
allowable parameter space for CRs and magnetic fields. Motivated by immanent 
challenges to explain the extent of giant radio halos within the hadronic model, 
we suggest a new hybrid leptonic/hadronic scenario meant to unify the apparently 
distinct classes of giant radio halos, mini halos, and steep spectrum radio sources in 
Section~\ref{sec:discussion_hadronic}. In Section~\ref{sec:4}, we apply our
extended hadronic model to a cosmologically complete mock galaxy cluster 
catalog built from the MultiDark $N$-body $\Lambda$CDM simulation
in \cite{paper1}, hereafter Paper~I. 
We compare the resulting modeled radio-to-X-ray and radio-to-SZ
scaling relations to current observations and show how they vary for different
choices of our CR and magnetic field parametrizations. In Section~\ref{sec:5}, we show
the radio luminosity functions, compare them to current observational
constraints, and provide predictions of the hadronically-generated
RHs for the LOFAR cluster survey. Finally, in Section~\ref{sec:6}, we present 
our conclusions.} In this work, the cluster mass
$M_{\Delta}$ and radius $R_{\Delta}$ are defined with respect to a density that
is $\Delta=200$ or $\Delta=500$ times the \emph{critical} density of the
Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$,
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of $H_0 = 100 \times
h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cosmic Ray Modeling}
\label{sec:2.3}
We assume a power-law for the spectral distribution of CR protons, $f(R,p) \dd
p=C(R) p^{-\alpha} \dd p$, which is the effective one-dimensional momentum
distribution (assuming isotropy in momentum space). {\bf We are interested 
in calculating the radio synchrotron and gamma-ray emission resulting from 
secondaries of hadronic CR interactions with protons of the ICM (see
Appendix~\ref{app:A} and Appendix~\ref{app:C} for the details).}

The spatial CR distribution $C(R)$ within a galaxy cluster is governed by an interplay
of CR advection, streaming, and diffusion. The advection of CRs by turbulent gas
motions is dominated by the largest eddy turnover time $\tau_{\rmn{tu}}\sim
L_{\rmn{tu}}/ \vel_{\rmn{tu}}$. Here, $L_{\rmn{tu}}$ denotes the turbulent
injection scale (typically of order the core radius) and $\vel_{\rmn{tu}}$ is
the associated turbulent velocity that approaches the sound speed
$\vel_{\rmn{s}}$ for transsonic turbulence after a cluster merger and relaxes to
small velocities afterwards. As a result of advection into the dense cluster
atmosphere, CRs are adiabatically compressed and experience a stratified
distribution in the cluster potential. The gradient of the CR number density
leads to a net CR streaming motion towards the cluster outskirts. Streaming CRs
excite Alfv{\'e}n waves on which they resonantly scatter
\citep{1969ApJ...156..445K}. This isotropizes the CRs' pitch angles, and thereby
reduces the CR bulk speed. Balancing the growth rate of the CR Alfv{\'e}n wave
instability with the wave damping rate due to non-linear Landau damping yields a
CR streaming speed of order the Alfv{\'e}n speed \citep{2001ApJ...553..198F}.
This increases considerably when balancing it with the turbulent damping rate,
which implies an inverse scaling with the CR number density \citep{2013arXiv1303.4746W}.
Once CR streaming depletes the CR number density, this causes a
run-away process with a rapidly increasing streaming speed that even surpasses
the sound speed because the smaller CR number density drives the CR Alfv{\'e}n
wave instability less efficiently. Hence, the crossing time of streaming CRs
over $L_{\rmn{tu}}$ is $\tau_{\rmn{st}}\sim \chi_B\,L_{\rmn{tu}}/
\vel_{\rmn{st}}$ with the streaming velocity given by $\vel_{\rmn{st}}\sim
\vel_{\rmn{s}}$ and $\chi_B\lesssim 1$ parametrizes the magnetic bending scale.
Magnetic bottlenecks for the macroscopic, diffusive CR transport, are critical
in lowering the microscopic streaming velocity of CR by some finite
factor. Therefore, we can define a turbulent propagation parameter
\begin{equation}
  \label{eq:gamma_tu}
  \gamma_{\rmn{tu}}\equiv\frac{\tau_{\rmn{st}}}{\tau_{\rmn{tu}}}=
  \frac{\chi_B\,\vel_{\rmn{tu}}}{\vel_{\rmn{st}}}
\end{equation}
that indicates the relative importance of advection versus CR streaming as the
dominant CR transport mechanism. After a merger, turbulent advective transport
should dominate yielding $\gamma_{\rmn{tu}}\gg 1$, which results in centrally enhanced
CR profiles. In contrast, in a relaxed cluster, CR streaming should be the
dominant transport mechanism implying $\gamma_{\rmn{tu}}\sim1$ and producing
flat CR profiles (for a detailed discussion of these processes, see
\citealp{2011A&A...527A..99E} and \citealp{2013arXiv1303.4746W}).

We propose here to take the spectral shape of the CR distribution function from
cosmological hydrodynamical simulation of clusters \citep{2010MNRAS.409..449P},
which however did not account for CR transport. {\bf To include the latter, we adapt 
the analytical formalism from \citet{2011A&A...527A..99E}. This yields a model that includes 
the necessary CR transport physics and is able to predict radio and gamma-ray 
emission. Note that this approach is not fully self-consistent and points to
the necessity of future hydrodynamical simulations to include the effect of CR
streaming and diffusion on the CR spectrum.}

@@

We will additionally study the CR-to-thermal
pressure $X_{\rmn{CR}}=P_{\rmn{CR}}/P_{\rmn{th}}$, where the CR pressure is
given by
\begin{equation}
  \label{eq:PCR}
  P_{\rmn{CR}}=\frac{g_{\rmn{CR}} C m_{\rmn{p}} c^{2}}{6}
  \sum_{i=1}^{3} \Delta_{i} \mathcal{B}_{1/(1+q^2)} \left(
    \frac{\alpha_{i}-1}{2},\frac{3-\alpha_{i}}{2} \right).
\end{equation}
Here, $c$ is the speed of light, $q=0.8$ is the low-momentum cutoff of the CR
distribution, and the normalization factors of the individual CR populations are
given by $\Delta_{i} = (0.767, 0.143, 0.0975)$ \citep[][see also
Appendix~\ref{app:A}]{2010MNRAS.409..449P}.

@@

>>>>

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{txfonts}
\usepackage[usenames,dvips]{color}
\newcommand\qc[1]{{\color{red} \bf #1}}

\newcommand{\CR}{\mathrm{CR}}
\newcommand{\expval}[1]{\left\langle #1 \right\rangle}
\newcommand{\RA}[3]{#1$^{\mathrm{h}}$#2$^{\mathrm{m}}$#3$^{\mathrm{s}}$}
\newcommand{\Dec}[3]{#1$^{\circ}$#2\arcmin#3\arcsec}
\newcommand{\dps}{\displaystyle}
\newcommand{\eps}{\varepsilon}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vel}{\upsilon}

@@

%MNRAS
\title[On the Physics of Radio Halos in Galaxy Clusters]{On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2}$, Christoph Pfrommer$^{3}$ and Francisco Prada$^{4,5,2}$\\
$^{1}$GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands, f.zandanel@uva.nl\\
$^{2}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany, christoph.pfrommer@h-its.org\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf The underlying physics of giant and mini radio halos in galaxy clusters
    is still an open question. We find that mini halos (such as in Perseus and
    Ophiuchus) can be explained by radio-emitting electrons that are generated
    in hadronic cosmic ray (CR) interactions with protons of the intra-cluster
    medium. By contrast, the hadronic model either fails to explain the extended
    emission of giant radio halos (as in Coma at low frequencies) or would
    require a flat CR profile, which can be realized through outward streaming
    and diffusion of CRs (in Coma and A2163 at 1.4~GHz).  We suggest that a
    second, leptonic component could be responsible for the missing flux in the
    outer parts of giant halos within a new hybrid scenario and we describe its
    possible observational consequences.  To study the hadronic emission
    component of the radio halo population statistically, we use a cosmological
    mock galaxy cluster catalog built from the MultiDark simulation. Because of
    the properties of CR streaming and the different scalings of the X-ray
    luminosity ($L_{\rmn{X}}$) and the Sunyaev-Zel'dovich flux ($Y$) with gas
    density, our model can simultaneously reproduce the observed bimodality of
    radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
    unimodal distribution of radio-halo luminosity versus $Y$; thereby
    suggesting a physical solution to this apparent contradiction. We predict
    radio halo emission down to the mass scale of galaxy groups, which
    highlights the unique prospects for low-frequency radio surveys (such as the
    LOFAR Tier 1 survey) to increase the number of detected radio halos by at
    least an order of magnitude.}
\end{abstract}

@@

{\bf The presence of large-scale diffuse radio synchrotron emission in clusters
  of galaxies proves the existence of relativistic electrons and magnetic fields
  permeating the intra-cluster medium (ICM). This diffuse cluster radio emission
  can be observationally classified into two phenomena: peripheral radio relics,
  which show irregular morphology and polarized emission and appear to trace
  merger and accretion shocks, as well as radio halos (see, e.g.,
  \citealp{2012A&ARv..20...54F}).  Radio \mbox{(mini-)}halos (RHs) are
  characterized by unpolarized radio emission, are centered on clusters and
  show a regular morphology, resembling the morphology of the thermal X-ray
  emission. However, the short cooling length of synchrotron-emitting electrons
  at GHz frequencies ($\lesssim 100$~kpc) challenges theoretical model to
  explain the large-scale radio emission that extends over several Mpcs and
  calls for an efficient in-site acceleration process of electrons.}

Two principal models have been proposed to explain RHs.  In the ``hadronic
model'' the radio emitting electrons are produced in hadronic cosmic ray (CR)
proton interactions with protons of the ambient thermal ICM, requiring only a
very modest fraction of (at most) a few percent of CR-to-thermal pressure
(\citealp{1980ApJ...239L..93D,1982AJ.....87.1266V, 1999APh....12..169B,
  2000A&A...362..151D, 2001ApJ...562..233M,2001ApJ...559...59M,
  2003MNRAS.342.1009M,2003A&A...407L..73P, 2004A&A...413...17P,
  2004MNRAS.352...76P, 2007IJMPA..22..681B, 2008MNRAS.385.1211P,
  2008MNRAS.385.1242P, 2009JCAP...09..024K, 2010MNRAS.401...47D,
  2010arXiv1003.0336D, 2010arXiv1003.1133K, 2010arXiv1011.0729K,
  2011A&A...527A..99E}).  CR protons and heavier nuclei, like electrons, can be
accelerated and injected into the ICM by structure formation shocks, active
galactic nuclei (AGN) and galactic winds.  Due to their higher masses with
respect to electrons, CR protons are accelerated more efficiently to
relativistic energies and are expected to show a ratio of the spectral energy
flux of CR protons to electrons above 1 GeV of about 100, similarly to what is
observed in our Galaxy \citep{2002cra..book.....S}. Additionally, CR protons
have a radiative cooling time larger than that of the electrons by the square of
the mass ratio and therefore can accumulate in clusters for cosmological times
\citep{1996SSRv...75..279V}. In contrast, CR electrons suffer more severe energy
losses via synchrotron and inverse Compton emission at particle energies $E
\gtrsim 100$~MeV, and Coulomb losses below that energy range.

@@

RHs can be divided in two classes. Giant radio halos are typically associated
with merging clusters and have large extensions, e.g., the Coma RH has an
extension of about 2~Mpc. Radio mini-halos are associated with relaxed clusters
that harbor a cool core and typically extend over a few hundred kilo-parsecs,
e.g., the Perseus radio mini-halo has an extension of about 0.2~Mpc.  The
observed morphological similarities with the thermal X-ray emission suggests
that RHs may be of hadronic origin. In fact, cool-core clusters (CCCs) are
characterized by high thermal X-ray emissivities and ICM densities that are more
peaked in comparison to non cool-core clusters (NCCCs) that often show
signatures of cluster mergers (e.g., \citealp{2008A&A...487..431C}). This
distinct difference in the ICM density structure of CCCs and NCCCs would be
reflected in the morphology of the two observed classes of RHs.

The RH luminosity seems to be correlated to the thermal X-ray luminosity (e.g.,
\citealp{2009A&A...507..661B,2011A&A...527A..99E}). However, a large fraction of
clusters does not exhibit significant diffuse synchrotron emission at current
sensitivity limits. Stacking subsamples of luminous X-ray clusters reveals a
signal of extended diffuse radio emission that is below the radio upper limits
on individual clusters \citep{2011ApJ...740L..28B} suggesting that at least a
subset of apparently ``radio-quiet'' clusters shows a low-level diffuse
emission. Galaxy clusters with the same thermal X-ray luminosity show an
apparent bimodality with respect to their radio luminosity.  This suggests the
existence of a switch-on/switch-off mechanism that is able to change the radio
luminosity by more than one order of magnitude.  While such a mechanism could be
easily realized in the framework of the re-acceleration model
\citep{2009A&A...507..661B}, the \emph{classical} hadronic model predicts the
presence of RHs in all clusters. The failure to reproduce the observed cluster
radio-to-X-ray bimodality was one of the main criticisms against the hadronic
model.  Additionally, the classical hadronic model cannot reproduce some
spectral features observed in clusters, such as the total spectral (convex)
curvature claimed in the Coma RH or the spectral steepening observed at the
boundary of some RHs.  However, the recent report of spectral flattening with
frequency of the RH in A2256 \citep{2012A&A...543A..43V} could easily be
accommodated in the hadronic model, which naturally produces such a concave
spectrum \citep{2010MNRAS.409..449P}.  This raises the interesting question
whether such a variability among different sources that are generally classified
as ``radio halos'' signals the presence of richer underlying physics---a question
that we will address in this paper.

\cite{2011A&A...527A..99E} tried to asses these problems of the classical
hadronic model by analyzing CR transport processes within a cluster. While CR
advection with turbulently driven flows results in centrally enhanced CR
profiles, the propagation in form of CR streaming and diffusion produces a
flattening of CR profiles. Hence, different CR transport phenomena may also
account for the observed bimodality of the radio luminosity in the hadronic
model and may also explain the spectral features observed in some clusters. {\bf
  This has been recently confirmed by \cite{2013arXiv1303.4746W}. In particular,
  by considering turbulent damping, they show that CRs can stream at
  super-Alfv{\'e}nic velocities.}  Note that these phenomena were not considered in
earlier analytical works (e.g., \citealp{2004A&A...413...17P}) as well as in
previous hydrodynamic simulations (e.g., \citealp{2001ApJ...562..233M,
  2008MNRAS.385.1211P, 2010MNRAS.409..449P}).  {\bf A satisfactory theory of CR
  transport in clusters does not yet exist. However, CR streaming and diffusion
  may represent an intriguing solution for the issues of the classical hadronic
  model.}

@@

An observational test that is able to disentangle between the hadronic and
re-acceleration models is the gamma-ray emission resulting from neutral pion
decays, a secondary product of the hadronic CR interaction with protons of the
ICM, which is not predicted by the re-acceleration model. Such observational
efforts have been undertaken in the last few years (for space-based cluster
observations in the GeV-band, see \citealt{2003ApJ...588..155R,
  2010JCAP...05..025A,2010ApJ...717L..71A,2012arXiv1207.6749H,
  2013arXiv1308.5654T,2013arXiv1308.6278H,2013arXiv1309.0197P}; for ground-based
observations in the energy band $\gtrsim100$~GeV, see
\citealt{2006ApJ...644..148P, 2008AIPC.1085..569P,
  2009arXiv0907.0727T,2009A&A...495...27A, 2009arXiv0907.3001D,
  2009arXiv0907.5000G, cangaroo_clusters, 2009ApJ...706L.275A,
  2010ApJ...710..634A, 2011arXiv1111.5544M,2012...VERITAS,2012A&A...545A.103H})
without being able to detect cluster gamma-ray emission. Current gamma-ray
limits enable us to constrain the average CR-to-thermal pressure to be less than
a couple percent, and the maximum CR acceleration efficiency at structure
formation shocks to be $<50$ per cent. Alternatively, this may indeed suggest the
presence of non-negligible CR transport processes into the outer cluster
regions.

@@

{\bf The main scope of this work is to account for CR transport processes in the
  classical hadronic model and to provide forecasts for future radio
  surveys. The outline is as follows. In Section~\ref{sec:2.3}, we construct a
  model for the CR proton distribution in clusters that merges results of
  hydrodynamic cluster simulations and an analytical model for microscopic CR
  transport processes. In Section~\ref{sec:3}, we model observed surface
  brightness profiles of individual RHs within the hadronic scenario and explore
  the allowable parameter space for CRs and magnetic fields. Motivated by
  immanent challenges to explain the extent of giant radio halos within the
  hadronic model, we suggest a new hybrid leptonic/hadronic scenario meant to
  unify the apparently distinct classes of giant radio halos, mini halos, and
  steep spectrum radio sources in Section~\ref{sec:discussion_hadronic}. In
  Section~\ref{sec:4}, we apply our extended hadronic model to a cosmologically
  complete mock galaxy cluster catalog built from the MultiDark $N$-body
  $\Lambda$CDM simulation in \cite{paper1}, hereafter Paper~I.  We compare the
  resulting modeled radio-to-X-ray and radio-to-SZ scaling relations to current
  observations and show how they vary for different choices of our CR and
  magnetic field parametrizations. In Section~\ref{sec:5}, we show the radio
  luminosity functions, compare them to current observational constraints, and
  provide predictions of the hadronically-generated RHs for the LOFAR cluster
  survey. Finally, in Section~\ref{sec:6}, we present our conclusions.} In this
work, the cluster mass $M_{\Delta}$ and radius $R_{\Delta}$ are defined with
respect to a density that is $\Delta=200$ or $500$ times the
\emph{critical} density of the Universe. We adopt density parameters of
$\Omega_{\rmn{m}}=0.3$, $\Omega_{\Lambda}=0.7$ and today's Hubble constant of
$H_0 = 100 \, h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cosmic Ray Modeling}
\label{sec:2.3}
We assume a power-law for the spectral distribution of CR protons, $f(R,p) \dd
p=C(R) p^{-\alpha} \dd p$, which is the effective one-dimensional momentum
distribution (assuming isotropy in momentum space). The spatial CR distribution
$C(R)$ within a galaxy cluster is governed by an interplay of CR advection,
streaming, and diffusion. The advection of CRs by turbulent gas motions is
dominated by the largest eddy turnover time $\tau_{\rmn{tu}}\sim L_{\rmn{tu}}/
\vel_{\rmn{tu}}$. Here, $L_{\rmn{tu}}$ denotes the turbulent injection scale
(typically of order the core radius) and $\vel_{\rmn{tu}}$ is the associated
turbulent velocity that approaches the sound speed $\vel_{\rmn{s}}$ for
transsonic turbulence after a cluster merger and relaxes to small velocities
afterwards. As a result of advection into the dense cluster atmosphere, CRs are
adiabatically compressed and experience a stratified distribution in the cluster
potential. The gradient of the CR number density leads to a net CR streaming
motion towards the cluster outskirts. Streaming CRs excite Alfv{\'e}n waves on
which they resonantly scatter \citep{1969ApJ...156..445K}. This isotropizes the
CRs' pitch angles, and thereby reduces the CR bulk speed. Balancing the growth
rate of the CR Alfv{\'e}n wave instability with the wave damping rate due to
non-linear Landau damping yields a CR streaming speed of order the Alfv{\'e}n
speed \citep{2001ApJ...553..198F}.  This increases considerably when balancing
it with the turbulent damping rate, which implies an inverse scaling with the CR
number density \citep{2013arXiv1303.4746W}.  Once CR streaming depletes the CR
number density, this causes a run-away process with a rapidly increasing
streaming speed that even surpasses the sound speed because the smaller CR
number density drives the CR Alfv{\'e}n wave instability less
efficiently. Hence, the crossing time of streaming CRs over $L_{\rmn{tu}}$ is
$\tau_{\rmn{st}}\sim \chi_B\,L_{\rmn{tu}}/ \vel_{\rmn{st}}$ with the streaming
velocity given by $\vel_{\rmn{st}}\sim \vel_{\rmn{s}}$ and $\chi_B\lesssim 1$
parametrizes the magnetic bending scale.  Magnetic bottlenecks for the
macroscopic, diffusive CR transport, are critical in lowering the microscopic
streaming velocity of CR by some finite factor. Therefore, we can define a
turbulent propagation parameter
\begin{equation}
  \label{eq:gamma_tu}
  \gamma_{\rmn{tu}}\equiv\frac{\tau_{\rmn{st}}}{\tau_{\rmn{tu}}}=
  \frac{\chi_B\,\vel_{\rmn{tu}}}{\vel_{\rmn{st}}}
\end{equation}
that indicates the relative importance of advection versus CR streaming as the
dominant CR transport mechanism. After a merger, turbulent advective transport
should dominate yielding $\gamma_{\rmn{tu}}\gg 1$, which results in centrally enhanced
CR profiles. In contrast, in a relaxed cluster, CR streaming should be the
dominant transport mechanism implying $\gamma_{\rmn{tu}}\sim1$ and producing
flat CR profiles (for a detailed discussion of these processes, see
\citealp{2011A&A...527A..99E} and \citealp{2013arXiv1303.4746W}).

We propose here to take the spectral shape of the CR distribution function from
cosmological hydrodynamical simulation of clusters \citep{2010MNRAS.409..449P},
which however did not account for CR transport. {\bf To include the latter, we
  adapt the analytical formalism of \citet{2011A&A...527A..99E}. This results in
  a model that includes the necessary CR transport physics and is able to
  predict radio and gamma-ray emission. Note that this approach is not fully
  self-consistent and points to the necessity of future hydrodynamical
  simulations to include the effect of CR streaming and diffusion on the CR
  spectrum.}

@@

@@

Coma story:
